---
title: "nearspectRa"
author: "Methun George & Dr. Steffen Neumann"
date: "2025-02-14"
output: 
  html_document :
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This vignette serves as a user guide for 'nearspectRa,' demonstrating the package's functions with code examples and expected outputs.

# SummarizedExperiment for ".asd" files

```{r}
# Function name : read_summarizedexperiment.asd

# Read ASD Spectral Files and Create a SummarizedExperiment Object

source("/Users/methungeorge/Desktop/nearspectRa/R/read_summarizedexperiment_asd.R")

# Example Usage 

# Define the path to your asd file or directory containing the files
path <- "/Users/methungeorge/Desktop/IPB/test/data"

# Save the output of function 
se_asd <- read_summarizedexperiment_asd(path)

# To see the output
print(se_asd)
```

The features used in this analysis are the wavelengths of the NIRS spectra.  Metadata associated with each wavelength is stored in the row names, and sample data are represented in the columns.  A summary of the SummarizedExperiment dimensions is provided at the top.


```{r}
# To access the column names of the SummarizedExperiment
colnames(se_asd)
```


```{r}
# To access the row names of the SummarizedExperiment
head(rownames(se_asd))
```


```{r}
# To access the assay data
assay_data <- assay(se_asd, "counts")
head(assay_data)

```

# SummarizedExperiment for ".sig" files

```{r}
# Function name : read_summarizedexperiment.sig

# Read sig Spectral Files and Create a SummarizedExperiment Object

source("/Users/methungeorge/Desktop/nearspectRa/R/read_summarizedexperiment_sig.R")
path <- "/Users/methungeorge/Desktop/IPB/test/sig_data"
se_sig <- read_summarizedexperiment_sig(path)

# To see the output 
print(se_sig)
```


```{r}
# To access the column names of the SummarizedExperiment
colnames(se_sig)
```

```{r}
# To access the row names of the SummarizedExperiment
head(rownames(se_sig))
```

```{r}
# To access the assay data
assay_data_sig <- assay(se_sig, "counts")
head(assay_data_sig)
```


# Visualization

```{r}
library(ComplexHeatmap)

Heatmap(
  assay(se_asd, "counts"),
  name = "Reflectance",
  column_names_gp = gpar(fontsize = 10),
  row_names_gp = gpar(fontsize = 8)
)



```





